---
title: Methods for Risk Evaluation \#2
author: Jingchen (Monika) Hu 
institute: Vassar College
date: Data Confidentiality
output:
  beamer_presentation:
    includes:
      in_header: ../LectureStyle.tex
slide_level: 2
fontsize: 11pt

---


```{r setup, include=FALSE}
require(dplyr)
require(ggplot2)
knitr::opts_chunk$set(echo = TRUE)
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
```

## Outline

\tableofcontents[hideallsubsections]


# Categorical example \#2: synthetic ACS samples

## Recap of $m = 1$

- A good practice to write functions to calculate various quantities

- When ```m > 1```
    - create ```c_vector, T_vector, K_vector, F_vector``` as matrices ($n-by-m$)
    - create ```s``` as a vector of length $m$
    - add nested loops when necessary
    - create ```exp_match_risk, true_match_rate, false_match_rate``` as vectors of length $m$
    - ```syndata``` is a list
    
## ACS sample information

\scriptsize

| Variable | Information                                  |
| :------------ | :---------------------------------------------------- |
| SEX           | 1 = male, 2 = female                                  |
| RACE          | 1 = White alone, 2 = Black or African American alone, 3 =                      American Indian alone, 4 = other, 5 = two or more races, 6 =                   Asian alone                                           |
| MAR           | 1 = married, 2 = widowed, 3 = divorced, 4 = separated, 5 = never                married                                               |
| LANX          | 1 = speaks another language, 2 = speaks only English  |
| WAOB          | born in: 1 = US state, 2 = Puerto Rico and US island areas, oceania and                 at sea, 3 = Latin America, 4 = Asia, 5 = Europe, 6 = Africa, 7 =                Northern America                                      |
| DIS           | 1 = has a disability, 2 = no disability               |
| HICOV         | 1 = has health insurance coverage, 2 = no coverage    |
| MIG           | 1 = live in the same house (non movers), 2 = move to outside US                and Puerto Rico, 3 = move to different house in US or Puerto Rico       |
| SCH           | 1 = has not attended school in the last 3 months, 2 = in public                school or college, 3 = in private school or college or home                    school                                                |

## ACS sample information cont'd

- ```ACSdata_org```: the original ACS sample

- ```ACSdata_syn```, ```ACSdata_syn2```, and ```ACSdata_syn3```: three synthetic ACS samples
    - four variables are synthesized: ```LANX, WAOB, DIS, HICOV```
    - ```m = 3```

\pause

- Known variables: ```SEX, RACE, MAR```

- Goal: use this information to identify records in ```ACSdata_syn```, obtain the 3 summaries

```{r message=FALSE, size = "footnotesize"}
ACSdata_org <- read.csv(file = "ACSdata_org.csv")
ACSdata_syn <- read.csv(file = "ACSdata_syn.csv")
ACSdata_syn2 <- read.csv(file = "ACSdata_syn2.csv")
ACSdata_syn3 <- read.csv(file = "ACSdata_syn3.csv")
ACSdata_syn_all <- list(ACSdata_syn, ACSdata_syn2, ACSdata_syn3)
```

## The IdentificationRisk function - page 1

```{r message=FALSE, size = "footnotesize", eval = FALSE}
IdentificationRisk <- function(origdata, syndata, known.vars, syn.vars, 
                               m, n){
  origdata <- origdata
  syndata <- syndata
  m <- m 
  n <- n 

  c_vector <- matrix(rep(NA, n*m), ncol = m)
  T_vector <- matrix(rep(NA, n*m), ncol = m)

```

## The IdentificationRisk function - page 2

```{r message=FALSE, size = "footnotesize", eval = FALSE}
  for (i in 1:n){
    for (k in 1:m){
      syndata_k <- syndata[[k]]
      match_k <- (eval(parse(text=paste("origdata$",syn.vars,"[i]==
                                      syndata_k$",syn.vars,sep="",
                                      collapse="&")))&
                  eval(parse(text=paste("origdata$",known.vars,"[i]==
                                        syndata_k$",known.vars,sep="",
                                        collapse="&"))))
      match.prob_k <- ifelse(match_k,1/sum(match_k),0)
      
      if (max(match.prob_k) > 0){
        c_vector[i, k] <- length(match.prob_k[match.prob_k 
                                              == max(match.prob_k)])
      }
      else
        c_vector[i, k] <- 0
      T_vector[i, k] <- is.element(i,rownames(origdata)
                                   [match.prob_k == max(match.prob_k)])
    }
  }
```

## The IdentificationRisk function - page 3

```{r message=FALSE, size = "footnotesize", eval = FALSE}
  K_vector <- matrix(rep(NA, n*m), ncol = m)
  F_vector <- matrix(rep(NA, n*m), ncol = m)
  for (k in 1:m){
    K_vector[, k] <- (c_vector[, k]*T_vector[, k]==1)
    F_vector[, k] <- (c_vector[, k]*(1 - T_vector[, k])==1)
  }
  
  s_vector <- rep(NA, m)
  exp_match_risk_vector <- rep(NA, m)
  true_match_rate_vector <- rep(NA, m)
  false_match_rate_vector <- rep(NA, m)

  for (k in 1:m){
    s_vector[k] <- length(c_vector[c_vector[, k]==1 & 
                                     is.na(c_vector[, k])==FALSE, k])
    nonzero_c_index <- which(c_vector[, k]>0)
    exp_match_risk_vector[k] <- sum(1/c_vector[nonzero_c_index, k] 
                              * T_vector[nonzero_c_index, k])
    true_match_rate_vector[k] <- sum(na.omit(K_vector[, k]))/n
    false_match_rate_vector[k] <- sum(na.omit(F_vector[, k]))/s_vector[k]
  }
```

## The IdentificationRisk function - page 4

```{r message=FALSE, size = "footnotesize", eval = FALSE}
  res_r <- list(s_vector = s_vector,
                exp_match_risk_vector = exp_match_risk_vector,
                true_match_rate_vector = true_match_rate_vector,
                false_match_rate_vector = false_match_rate_vector,
                c_vector = c_vector,
                T_vector = T_vector,
                K_vector = K_vector,
                F_vector = F_vector
  )
  return(res_r)
}
```

```{r message=FALSE, echo = FALSE}
IdentificationRisk <- function(origdata, syndata, known.vars, syn.vars, m, n){
  origdata <- origdata
  syndata <- syndata
  m <- m 
  n <- n  

  c_vector <- matrix(rep(NA, n*m), ncol = m)
  T_vector <- matrix(rep(NA, n*m), ncol = m)

  for (i in 1:n){
    for (k in 1:m){
      syndata_k <- syndata[[k]]
      match_k <- (eval(parse(text=paste("origdata$",syn.vars,"[i]==
                                      syndata_k$",syn.vars,sep="",collapse="&")))&
                  eval(parse(text=paste("origdata$",known.vars,"[i]==
                                        syndata_k$",known.vars,sep="",collapse="&"))))
      match.prob_k <- ifelse(match_k,1/sum(match_k),0)
      if (max(match.prob_k) > 0){
        c_vector[i, k] <- length(match.prob_k[match.prob_k == max(match.prob_k)])
      }
      else
        c_vector[i, k] <- 0
      T_vector[i, k] <- is.element(i,rownames(origdata)[match.prob_k==max(match.prob_k)])
    }
  }
  K_vector = matrix(rep(0, n*m), ncol = m)
  F_vector = matrix(rep(0, n*m), ncol = m)
  for (k in 1:m){
    K_vector[, k] = (c_vector[, k]*T_vector[, k]==1)
    F_vector[, k] = (c_vector[, k]*(1 - T_vector[, k])==1)
  }

  s_vector <- rep(NA, m)
  exp_match_risk_vector <- rep(NA, m)
  true_match_rate_vector <- rep(NA, m)
  false_match_rate_vector <- rep(NA, m)

  for (k in 1:m){
    s_vector[k] <- length(c_vector[c_vector[, k]==1&is.na(c_vector[, k])==FALSE, k])
    nonzero_c_index = which(c_vector[, k]>0)
    exp_match_risk_vector[k] <- sum(1/c_vector[nonzero_c_index, k]*T_vector[nonzero_c_index, k])
    true_match_rate_vector[k] <- sum(na.omit(K_vector[, k]))/n
    false_match_rate_vector[k] <- sum(na.omit(F_vector[, k]))/s_vector[k]
  }

  res_r <- list(s_vector = s_vector,
                exp_match_risk_vector = exp_match_risk_vector,
                true_match_rate_vector = true_match_rate_vector,
                false_match_rate_vector = false_match_rate_vector,
                c_vector = c_vector,
                T_vector = T_vector,
                K_vector = K_vector,
                F_vector = F_vector
  )
  return(res_r)
}
```

## Running the function

```{r message=FALSE, size = "footnotesize"}
known.vars <- c("SEX", "RACE", "MAR") 
syn.vars <- c("LANX", "WAOB", "DIS", "HICOV") 
n <- dim(ACSdata_org)[1]
m <- length(ACSdata_syn_all)

output <- IdentificationRisk(ACSdata_org, ACSdata_syn_all, 
                             known.vars, syn.vars, m, n)
```

## Results and discussion

```{r message=FALSE, size = "footnotesize"}
mean(output[["exp_match_risk_vector"]])
mean(output[["true_match_rate_vector"]])
mean(output[["false_match_rate_vector"]])
```

## Results and discussion

- The 41.47 expected match risk: $\frac{41.47}{10000} = 0.000042$ probability on average for each record to be correctly identified

\pause

- The 0.0006 true match rate: $0.0006 \times 10000 = 6$ records are correct unique matches

\pause

- The 0.96 false match rate: among the 161 (the value of $s$) unique matches, about 155 are false matches, i.e. they are not the true matches

```{r message=FALSE, size = "footnotesize"}
mean(output[["s_vector"]])
```
\pause 

- Overall, the identification disclosure risks for the synthetic ACS sample seem very low, indicating a high level of confidentiality protection of the synthetic ACS data


# Continuous example: synthetic CE sample

## What do we need?

- $c_i$: the number of records with the highest match probability for the target record $i$.
    1. the records with the highest match probability for record $i$ are a subset of all the records sharing the same (or very similar) known variables and synthesized variables.
    2. e.g. for categorical variable(s), we can consider all records in the same known variables and synthesized variables with record $i$.
    3. e.g. for continuous variable(s), we can consider all records very similar known variables and synthesized variables from record $i$, for example, within a certain distance in terms of the continuous variable(s).
    
- $T_i$: if the true match is among the $c_i$ units, $T_i = 1$; otherwise $T_i = 0$.

## What do we need? cont'd

- $K_i$: if the true match is the unique match (i.e. $c_i T_i = 1$), $K_i = 1$; otherwise $K_i = 0$.

- $F_i$: if there is a unique match but it is not the true match (i.e. $c_i (1 - T_i) = 1$), $F_i = 1$; otherwise $F_i = 0$.

- $N$: the total number of target records; typically $N = n$, the number of records in the sample.

- $s$: the number of uniquely matched records (i.e. $\sum_{i=1}^{n} c_i = 1$).

## The three summaries

- The expected match risk
    - on average how likely it is to find the correct match for each record, and for the sample as a whole
    
\begin{equation}
\sum_{i=1}^{n} \frac{T_i}{c_i}
\end{equation}

## The three summaries con'td

- The true match rate
    - how large a percentange of true unique mathces exists

\begin{equation}
\sum_{i=1}^{n} \frac{K_i}{N}
\end{equation}

- The false match rate
    - the percentage of unique matches to be false matches

\begin{equation}
\sum_{i=1}^{n} \frac{F_i}{s}
\end{equation}

## What are your methods?

## Some coding techniques

- Categorical case:

```{r message=FALSE, size = "footnotesize", eval = FALSE}
match_k <- (eval(parse(text=paste("origdata$",syn.vars,"[i]==
                                      syndata_k$",syn.vars,sep="",
                                      collapse="&")))&
                  eval(parse(text=paste("origdata$",known.vars,"[i]==
                                        syndata_k$",known.vars,sep="",
                                        collapse="&"))))
```

## Some coding techniques cont'd

- Continuous case:
    - example of synthesized variables: one univariate continuous, e.g. (```syn.vars <- c("Income")```)
    - ```radius```: the distance from the true value of the synthesized univariate continuous value, e.g. ```Income```
    - the distance can be an absolute value (e.g. \$500 for every CU) or a percentage (e.g. 20\% for every CU)
    
```{r message=FALSE, size = "footnotesize", eval = FALSE}
match_k<-(eval(parse(text=paste("origdata$",known.vars,"[i]==
                                        syndata_k$",known.vars,sep="",
                                        collapse="&")))&
                  (eval(parse(text=paste("syndata_k$",syn.vars,"<=
                                      origdata$",syn.vars,"[i]+",
                                      radius,sep="",collapse="&")))&
                     eval(parse(text=paste("syndata_k$",syn.vars,">=
                                      origdata$",syn.vars,"[i]-",
                                      radius,sep="",collapse="&")))))
```

## Discussions

- What if we have more than one synthesized continuous variables? How can you create the corresponding ```radius```?