---
title: "Assignment #03"
author: "Yitong Wu"
date: "February 18, 2020"
output: html_notebook
---

(a) Use your own synthesis model to synthesize m = 1 synthetic dataset for the CE sample.
```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(fastDummies)
knitr::opts_chunk$set(echo = TRUE)
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
require(runjags)
require(coda)
```

```{r}
data <- read.csv("CEdata.csv")
data$Rural = fastDummies::dummy_cols(data$UrbanRural)[,names(fastDummies::dummy_cols(data$UrbanRural)) == ".data_2"]
data$Race_Black = fastDummies::dummy_cols(data$Race)[,names(fastDummies::dummy_cols(data$Race)) == ".data_2"]
data$Race_NA = fastDummies::dummy_cols(data$Race)[,names(fastDummies::dummy_cols(data$Race)) == ".data_3"]
data$Race_Asian = fastDummies::dummy_cols(data$Race)[,names(fastDummies::dummy_cols(data$Race)) == ".data_4"]
data$Race_PI = fastDummies::dummy_cols(data$Race)[,names(fastDummies::dummy_cols(data$Race)) == ".data_5"]
data$Race_M = fastDummies::dummy_cols(data$Race)[,names(fastDummies::dummy_cols(data$Race)) == ".data_6"]
data$logInc <- log(data$Income)
```

```{r message = FALSE, size = "footnotesize"}
modelString <-"
model {
## sampling
for (i in 1:N){
y[i] ~ dnorm(beta0 + beta1*x_rural[i] +
beta2*x_race_B[i] + beta3*x_race_N[i] +
beta4*x_race_A[i] + beta5*x_race_P[i] + beta6*x_race_M[i], invsigma2)
}
## priors
beta0 ~ dnorm(mu0, g0)
beta1 ~ dnorm(mu1, g1)
beta2 ~ dnorm(mu2, g2)
beta3 ~ dnorm(mu3, g3)
beta4 ~ dnorm(mu4, g4)
beta5 ~ dnorm(mu5, g5)
beta6 ~ dnorm(mu6, g6)
invsigma2 ~ dgamma(a, b)
sigma <- sqrt(pow(invsigma2, -1))
}
"
```

```{r}
y = as.vector(data$logInc)
x_rural = as.vector(data$Rural)
x_race_B = as.vector(data$Race_Black)
x_race_N = as.vector(data$Race_NA)
x_race_A = as.vector(data$Race_Asian)
x_race_P = as.vector(data$Race_PI)
x_race_M = as.vector(data$Race_M)
N = length(y)  
```

```{r message = FALSE, size = "footnotesize"}
the_data <- list("y" = y, 
                 "x_rural" = x_rural, "x_race_B" = x_race_B,
                 "x_race_N" = x_race_N, "x_race_A" = x_race_A,
                 "x_race_P" = x_race_P, "x_race_M" = x_race_M,
                 "N" = N,
                 "mu0" = 0, "g0" = 1, "mu1" = 0, "g1" = 1,
                 "mu2" = 0, "g2" = 1, "mu3" = 0, "g3" = 1,
                 "mu4" = 0, "g4" = 1, "mu5" = 0, "g5" = 1,
                 "mu6" = 0, "g6" = 1, 
                 "a" = 1, "b" = 1)
```

```{r message = FALSE, size = "footnotesize"}
initsfunction <- function(chain){
  .RNG.seed <- c(1,2)[chain]
  .RNG.name <- c("base::Super-Duper",
                 "base::Wichmann-Hill")[chain]
  return(list(.RNG.seed=.RNG.seed,
              .RNG.name=.RNG.name))
}
```

```{r message = FALSE, size = "footnotesize", results = 'hide'}
posterior_MLR <- run.jags(modelString,
                      n.chains = 1,
                      data = the_data,
                      monitor = c("beta0", "beta1", "beta2",
                                  "beta3", "beta4", "beta5",
                                  "beta6", "sigma"),
                      adapt = 1000,
                      burnin = 5000,
                      sample = 5000,
                      thin = 1,
                      inits = initsfunction)
```

```{r message=FALSE, size = "footnotesize"}
post <- as.mcmc(posterior_MLR)
```

```{r message=FALSE, size = "footnotesize"}
synthesize <- function(X_rural, X_RB, X_RN, X_RA, X_RP, X_RM, index, n){
  mean_Y <- post[index, "beta0"] +  X_rural * post[index, "beta1"] + X_RB * post[index, "beta2"] + X_RN * post[index, "beta3"] + X_RA * post[index, "beta4"] + X_RP * post[index, "beta5"] + X_RM * post[index, "beta6"]
  synthetic_Y <- rnorm(n, mean_Y, post[index, "sigma"])
  data.frame(synthetic_Y)
}
```

```{r}
n <- dim(data)[1]
Syndata <- synthesize(data$Rural, data$Race_Black, data$Race_NA, data$Race_Asian, data$Race_PI, data$Race_M, 1, n)
names(Syndata) <- c("SynlogInc")
```

(b) Make a scatter plot of the synthesized log(Income) against the original log(Income), and see what you find.
```{r}
Synthesis <- cbind(data$logInc, Syndata$SynlogInc)
Synthesis <- data.frame(Synthesis)
names(Synthesis) <- c("Orig", "Syn")
```

```{r}
ggplot(Synthesis, aes(x=Orig, y=Syn)) + geom_point()
```

(b) Compare the mean and median of log(Income), in the original dataset and the confidential dataset. Are they close to each other?
```{r}
mean(Synthesis$Orig)
median(Synthesis$Orig)
mean(Synthesis$Syn)
median(Synthesis$Syn)
```

(c) Compare the point estimate of the regression coefficients of log(Income) on log(Expenditure), in the original dataset and the confidential dataset. Are they close to each other?
```{r}
Synthesis <- cbind(Synthesis, log(data$Expenditure))
names(Synthesis) <- c("Orig", "Syn", "Ex")
model_orig <- lm(Orig ~ Ex, data=Synthesis)
model_syn <- lm(Syn ~ Ex, data=Synthesis)
summary(model_orig)
summary(model_syn)
```

(d) Evaluate the global utility measures of your synthesized log(Income) from your Bayesian synthesis model for the CE sample.

I. Propensity Score Measure
```{r}
original_data <- data.frame(data$logInc)
original_data$T <- 0
names(original_data) <- c("logInc", "T")
synthesis_data <- data.frame(Synthesis$Syn)
synthesis_data$T <- 1
names(synthesis_data) <- c("logInc", "T")
psm_data <- bind_rows(original_data, synthesis_data)
psm_data$logInc <- round(psm_data$logInc, 1)
psm_data$psm <- 0
1+N/2
```

```{r}
N <- length(psm_data$T)
test <- psm_data %>% filter(T==1)
for (i in 1:N) {
  a <- psm_data$logInc[i]
  score <- sum(test$logInc==a)
  psm_data$psm[i] <- score
}
psm_data$pscore <- psm_data$psm/994
```

```{r}
psm_ori <- psm_data %>% filter(T==0)
psm_syn <- psm_data %>% filter(T==1)
```

```{r}
ggplot(psm_ori, aes(x=pscore)) + geom_histogram(binwidth=.001)
ggplot(psm_syn, aes(x=pscore)) + geom_histogram(binwidth=.001)
```

