---
title: "M3SynthesisEval"
author: "Sarah Boese"
date: "3/24/2020"
output: pdf_document
---

```{r, message = FALSE}
library(ProbBayes)
library(dplyr)
library(ggplot2)
require(gridExtra)
library(reshape)
library(runjags)
library(coda)
library(tidyverse)
library(fastDummies)
```

Here I read in the original and synthetic ACS data. 
```{r}
ACSdata_org <- read.csv("ACSdata_org.csv")
ACSdata_syn_1 <- read.csv("ACSdata_syn.csv")
ACSdata_syn_2 <- read.csv("ACSdata_syn2.csv")
ACSdata_syn_3 <- read.csv("ACSdata_syn3.csv")
```

Now I create a vector of c_i's. I couldn't figure out a vectorized function so I would love to see an example of how to do so. 


Here I create a dataframe containing values for $c_i$ and $T_i$. Here I assume $T_i$ is one since all the known variables are the same in both synthetic and orginal dataframes. I'm not sure if that interpretation is correct. I apply vectorized functions from dplyr to calculate $K_i$ and $F_i$. From there I was able to calculate expected match rate, true match rate and false match rate.
```{r}
eval_func<-function(data){
  N<- nrow(data)
  c_i<-vector(length=N)
  for(j in 1:N){
    vecy<- as.numeric(as.vector(data[j,c(1,2,3)]))
    c_i[j]=nrow(filter(data, SEX==vecy[1] & RACE==vecy[2] & MAR==vecy[3]))
  }

  ones <- integer(N) +1
  analysis_data<-data.frame(c_i, ones)
  names(analysis_data)<-c("c_i", "T_i")
  s<-nrow(filter(analysis_data, c_i==1))
  analysis_data_1 <- analysis_data %>%
    mutate(K_i=if_else(c_i*T_i==1, 1, 0)) %>%
    mutate(F_i=if_else(c_i*(1-T_i)==1, 1, 0)) %>%
    mutate(expect_i=T_i/c_i) %>%
    mutate(true_i=K_i/N) %>%
    mutate(false_i=F_i/s)

  expected_match<-sum(analysis_data_1$expect_i)
  expected_match

  true_match<-sum(analysis_data_1$true_i)
  true_match

  false_match<-sum(analysis_data_1$false_i)
  false_match
  
  out<- data.frame(expected_match, true_match, false_match)
  names(out)<-c("expected_match", "true_match", "false_match")
  return(out)
}
```


```{r, error=FALSE}
syn_eval_1<-eval_func(ACSdata_syn_1)
syn_eval_2<-eval_func(ACSdata_syn_2)
syn_eval_3<-eval_func(ACSdata_syn_3)
```

