---
title: Differentially Private Synthetic Tabular Data
author: Jingchen (Monika) Hu 
institute: Vassar College
date: Data Confidentiality
output:
  beamer_presentation:
    includes:
      in_header: ../LectureStyle.tex
slide_level: 2
fontsize: 11pt

---


```{r setup, include=FALSE}
require(dplyr)
require(ggplot2)
knitr::opts_chunk$set(echo = TRUE)
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
```

## Outline

\tableofcontents[hideallsubsections]

# Introduction

## Recap of differential privacy

- Definitions: database, query, output, sensitivity, privacy budge, and added noise

- Implications of key terms in differential privacy: the relationship between sensitivity ($\Delta f$), privacy budget ($\epsilon$), and added noise

\pause 

- The Laplace Mechanism
    - adds random noise according to $\epsilon-$differential privacy guarantee
    - the noise is drawn from a Laplace distribution centered at 0, with scale $\frac{\Delta f}{\epsilon}$
    
- DP properties: example queries to the confidential CE databse

## Recap of synthetic data

- A conjugate Bayesian model for categorical variables: Dirichlet-multinomial; now for contingency tables

1. Suppose we have a count vector ${{\bf{y}}}$ of length $I$, with total number of records $y.$ (the sum of ${{\bf{y}}}$). The multinomial sampling model follows:

\begin{equation}
{{\bf{y}}} \mid {{\boldsymbol{\theta}}} \sim \textrm{Multinomial}(y.; {{\boldsymbol{\theta}}}).
\end{equation}

\pause

2. A conjugate prior for ${\bf{\theta}}$ is Dirichlet:
\begin{equation}
{{\boldsymbol{\theta}}} \sim \textrm{Dirichlet}({{\boldsymbol{\alpha}}}). 
\end{equation}

\pause

3. Due to conjugacy, we come to a Dirichlet posterior for ${\bf{\theta}}$:

\begin{equation}
{{\boldsymbol{\theta}}} \mid {{\bf{y}}} \sim \textrm{Dirichlet}({{\bf{y}}} + {{\boldsymbol{\alpha}}}).
\end{equation}

## Overview

- Can we make the Dirichlet-multinomial synthesizer satisfy $\epsilon-$differential privacy?

- The original $\epsilon-$differential privacy definition: a mechanism $\mathcal{M}$ with domain $\mathbb{N}^{|\mathcal{X}|}$ is $\epsilon-$differentially private for all $\mathcal{S} \subseteq \textrm{Range}(\mathcal{M})$ and for all $\bf{x}, \bf{y} \in \mathbb{N}^{|\mathcal{X}|}$ such that $\delta({\bf{x}}, {\bf{y}}) = 1$:

\begin{equation}
\left|\ln\left(\frac{Pr[\mathcal{M}({\bf{x}}) \in \mathcal{S}]}{Pr[\mathcal{M}({\bf{y}}) \in \mathcal{S}]}\right)\right| \leq \epsilon.
\end{equation}

- What to do for synthetic tabular data?

## Overview cont'd

- Now in synthetic tabular data: 

Let ${\bf{y}}$ denote the true count vector of length $I$, and ${\bf{x}}$ denote another count vector with Hamming distance 1 from ${\bf{y}}$ ($\delta({{\bf{x}}}, {{\bf{y}}}) = 1$) and $\sum_{i=1}^{I}x_i = \sum_{i=1}^{I}y_i$. Let $y.^*$ denote an $\epsilon-$differentially private synthetic count vector, and $\boldsymbol{\theta}$ denote model parameters vector. In such setting, $\epsilon-$differential privacy requires

\begin{equation}
\left|\ln\left(\frac{p({\bf{y}}^* \mid {\bf{y}}, {\boldsymbol{\theta}})}{p({\bf{y}}^* \mid {\bf{x}}, {\boldsymbol{\theta}})}\right)\right| \leq \epsilon.
\end{equation}

\pause

- If we can the Dirichlet-multinomial synthesizer satisfy $\epsilon-$differential privacy, we produce \textcolor{red}{differentially private synthetic tabular data}

# $\epsilon-$differential privacy with Dirichlet-multinomial

## The procedure
Abowd and Vilhuber (2008) and Machanavajjhala et al. (2008)

- To generate a differentially private synthetic count vector ${{\bf{y}}}^*$ given $y.^* = y.$ (the total sum is fixed):

1. Sample ${{\boldsymbol{\theta}}}^*$ from

\begin{equation}
{{\boldsymbol{\theta}}} \mid {{\bf{y}}} \sim \textrm{Dirichlet}({{\bf{y}}} + {{\boldsymbol{\alpha}}}),
\end{equation}
where \textcolor{red}{$\textrm{min}(\alpha_i) \geq \frac{y.^*}{\exp(\epsilon) - 1}$}.

\pause

2. Sample ${{\bf{y}}}^*$ from

\begin{equation}
{{\bf{y}}}^* \mid {{\boldsymbol{\theta}}}^* \sim \textrm{Multinomial}(y.^*; {{\boldsymbol{\theta}}}^*),
\end{equation}
and the generated count vector ${{\bf{y}}}^*$ satisfies $\epsilon-$differential privacy.

## Why it works?

- The posterior predictive distribution is:

\begin{eqnarray}
p({{\bf{y}}}^* \mid {\bf{y}}, {\boldsymbol{\alpha}}) &=& \int p({\bf{y}}^* \mid {\boldsymbol{\theta}}, {\boldsymbol{\alpha}}) \times p({\boldsymbol{\theta}} \mid {\bf{y}}, {\boldsymbol{\alpha}}) d {\boldsymbol{\theta}} \nonumber \\
&=& \int \frac{y.^* !}{\prod_{i=1}^{I} y^*_i !} \times \prod_{i=1}^{I} \theta_i^{y^*_i} \times  \frac{\Gamma(\sum_{i=1}^{I}y_i + \alpha_i)}{\prod_{i=1}^{I} \Gamma(y_i + \alpha_i)} \times \prod_{i=1}^{I} \theta_i^{y_i + \alpha_i - 1} d {\boldsymbol{\theta}} \nonumber \\
&=& \frac{y.^* !}{\prod_{i=1}^{I} y^*_i !} \times \frac{\Gamma(\sum_{i=1}^{I}y_i + \alpha_i)}{\prod_{i=1}^{I} \Gamma(y_i + \alpha_i)} \times \frac{\prod_{i=1}^{I}\Gamma(y^*_i + y_i + \alpha_i)}{\Gamma(\sum_{i=1}^{I}y^*_i + y_i + \alpha_i)}. 
\end{eqnarray}

## Why it works? cont'd

- To satisfy $\epsilon-$differential privacy, we require 

\begin{equation}
\left|\log\left(\frac{p({\bf{y}}^* \mid {\bf{y}}, {\boldsymbol{\alpha}})}{p({\bf{y}}^* \mid {\bf{x}}, {\boldsymbol{\alpha}})}\right)\right| = \left| \log\left( \frac{\prod_{i=1}^{I} \Gamma(\alpha_i + x_i)}{\prod_{i=1}^{I}\Gamma(\alpha_i + y_i)} \times \frac{\prod_{i=1}^{I}\Gamma(y^*_i + \alpha_i + y_i)}{\prod_{i=1}^{I}\Gamma(y^*_i + \alpha_i + x_i)} \right) \right| \leq \epsilon
\end{equation}

- ${\bf{x}}$ has Hamming distance 1 from ${\bf{y}}$ (i.e. $\delta({{\bf{x}}}, {{\bf{y}}}) = 1$) 
- $\sum_{i=1}^{I}x_i = \sum_{i=1}^{I}y_i$ (total sum is fixed)


## Why it works? cont'd

- Assume the the only differences in ${\bf{x}}$ and ${\bf{y}}$ exist between the pairs $(x_i, x_{i'})$ and $(y_i, y_{i'})$

- Without loss of generality, assume $x_i = y_i - 1$ and $x_{i'} = y_{i'} + 1$

\begin{equation}
\frac{p({\bf{y}}^* \mid {\bf{y}}, {\boldsymbol{\alpha}})}{p({\bf{y}}^* \mid {\bf{x}}, {\boldsymbol{\alpha}})} = \frac{\alpha_i + y_i}{\alpha_{i'} + y_{i'} - 1} \times \frac{y^*_{i'} + \alpha_{i'} + y_{i'} - 1}{y^*_i + \alpha_i + y_i},
\end{equation}
where $y^*_i + y^*_{i'} \leq y.^*$.

## Why it works? cont'd

\begin{equation*}
\frac{p({\bf{y}}^* \mid {\bf{y}}, {\boldsymbol{\alpha}})}{p({\bf{y}}^* \mid {\bf{x}}, {\boldsymbol{\alpha}})} = \frac{\alpha_i + y_i}{\alpha_{i'} + y_{i'} - 1} \times \frac{y^*_{i'} + \alpha_{i'} + y_{i'} - 1}{y^*_i + \alpha_i + y_i}
\end{equation*}

- Maximized when $y_{i'} = 1, y^*_i = 0$ and $y^*_{i'} = z.$
- Minimized when $y_{i'} = 0, y^*_{i'} = 0$ and $y^*_i = z.$

\begin{equation}
\frac{\alpha_i}{y.^* + \alpha_i} \leq \frac{p({\bf{y}}^* \mid {\bf{y}}, {\boldsymbol{\alpha}})}{p({\bf{y}}^* \mid {\bf{x}}, {\boldsymbol{\alpha}})} \leq \frac{y.^* + \alpha_{i'}}{\alpha_{i'}}.
\end{equation}


## Why it works? cont'd

- Now to satisfy $\epsilon-$differential privacy where $\left|\log\left(\frac{p({\bf{y}}^* \mid {\bf{y}}, {\boldsymbol{\alpha}})}{p({\bf{y}}^* \mid {\bf{x}}, {\boldsymbol{\alpha}})}\right)\right| \leq \epsilon$, we require

\begin{equation}
\epsilon = \log \left(\frac{y.^* + \textrm{min}(\alpha_i)}{\textrm{min}(\alpha_i)} \right),
\end{equation}
which results in 
\begin{equation}
\textrm{min}(\alpha_i) \geq \frac{y.^*}{\exp(\epsilon) - 1}.
\end{equation}

# Example: differentially private synthetic CE count table

## CE data


| Variable Name | Variable information                                  |
| :------------ | :---------------------------------------------------- |
| UrbanRural    | Binary; the urban / rural status of CU: 1 = Urban, 2 = Rural. |
| Income        | Continuous; the amount of CU income bfore taxes in past 12 months. |
| \textcolor{red}{Race}          | Categorical; the race category of the reference person: 1 = White, 2 = Black, 3 =                        Native American, 4 = Asian, 5 = Pacific Islander, 6 = Multi-race.  |
| Expenditure   | Continuous; CU's total expenditures in last quarter.  |

Table: Variables used in the CE database. Data taken from the 2017 CE public use microdata samples. 

## The contingency table of Race categories

```{r message = FALSE, size = "footnotesize"}
require(readr)
CEdata <- read_csv("CEdata.csv")
Race_Count <- CEdata %>% count(Race)
Race_Count
```


## Step 1: calculate $\boldsymbol{\alpha}$

- Expression for $\boldsymbol{\alpha}$
\begin{equation*}
\textrm{min}(\alpha_i) \geq \frac{y.^*}{\exp(\epsilon) - 1}
\end{equation*}

- Use privacy budget $\epsilon = 5$

```{r message = FALSE, size = "footnotesize"}
epsilon <- 5

alpha_min <- sum(Race_Count$n)/(exp(epsilon) - 1)
alpha_min
```

```{r message = FALSE, size = "footnotesize"}
alpha_vector <- rep(alpha_min, dim(Race_Count)[1])
alpha_vector
```

## Step 2: sample $\boldsymbol{\theta}^*$


- ${\bf{y}}$ is the original count vector of ```Race``` categories.

```{r message = FALSE, size = "footnotesize"}
y_vector <- Race_Count$n
y_vector
```

- With the calcuated ```alpha_vector```, sample ${{\boldsymbol{\theta}}^*}$ from $\textrm{Dirichlet}({\bf{y}} + {\boldsymbol{\alpha}})$

```{r message = FALSE, size = "footnotesize"}
require(gtools)
set.seed(123)
theta_DPsyn <- rdirichlet(n = 1, alpha = y_vector + alpha_vector)
theta_DPsyn
```


## Step 3: sample $\boldsymbol{y}^*$

- With the sampled ```theta_DPsyn```, we can sample ${{\bf{y}}^*}$ from $\textrm{Multinomial}(y.^*; {{\boldsymbol{\theta}}}^*)$

```{r message = FALSE, size = "footnotesize"}
y_DPsyn <- rmultinom(n = 1, size = sum(y_vector), 
                     prob = theta_DPsyn)
```

## Step 3: sample $\boldsymbol{y}^*$ cont'd

- Put original and synthetic side-by-side

```{r message = FALSE, size = "footnotesize"}
y_DPsyn <- rmultinom(n = 1, size = sum(y_vector), 
                     prob = theta_DPsyn)
y_both <- data.frame(y_vector, y_DPsyn)
names(y_both) <- c("original", "DPsynthetic")
y_both
```

## Summary and discussion

- The choice of privacy budget $\epsilon$ has great influence on the resulted differentially private synthetic contingency table

- What are your thoughts?

\pause

- The higher the value of $\epsilon$, the higher the utility

\pause

- Other differentially private synthetic tabular data models:
    1. beta-binomial (McClure and Reiter, 2012)
    2. gamma-Poisson (Quick, 2019)

## References

- Abowd, J. M., and L. Vilhuber. (2008). How Protective Are Synthetic Data? Privacy in Statistical Databases, 239–46.

- Machanavajjhala, A., D. and Kifer, J. and Abowd, J.and Gehrke, and L. Vilhuber. (2008). Privacy: Theory Meets Practice on the Map. The IEEE 24th International Conference on Data Engineering, 277–86.

- McClure, D., and J. P. Reiter. (2012). Differential Privacy and Statistical Disclosure Risk Measures: An Investigation with Binary Synthetic Data. Transactions on Data Privacy 5: 535–52.

- Quick, H. (2019). Generaing Poisson-distributeed differentially private synthetic data. arXiv:1906.00455.

