---
title: "ProjectDataset"
output:
  pdf_document: default
  html_notebook: default
---

AP VoteCast is a survey of the American electorate conducted in all 50 states by NORC at the University of Chicago for The Associated Press and Fox News. The survey is funded by AP. The survey of 138,929 registered voters was conducted October 29 to November 6, 2018, concluding as polls closed on Election Day. Interviews were conducted via phone and web, with 11,059 completing by phone and 127,870 completing by web.

```{r include=FALSE}
library(readr)
library(dplyr)
library(ggplot2)
library(fastDummies)
knitr::opts_chunk$set(echo = TRUE)
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
require(runjags)
require(coda)
```

```{r}
data <- read.csv("dataset_CD.csv")
```

```{r}
data$RACETH5 <- recode(data$RACETH5, "(1) White" = 1, "(2) African American or Black" = 2, "(3) Latino or Hispanic" = 3, "(4) Asian" = 4, "(5) Other" = 5, "(99) DON’T KNOW/SKIPPED/REFUSED (VOL)" = 99)
data$EDUC <- recode(data$EDUC, "(1) High school or less" = 1, "(2) Some college/assoc. degree" = 2, "(3) College graduate" = 3, "(4) Postgraduate study" = 4, "(99) DON’T KNOW/SKIPPED/REFUSED (VOL)" = 99)
data$INCOME <- recode(data$INCOME, "(1) Under $25,000" = 1, "(2) $25,000-$49,999" = 2, "(3) $50,000-$74,999" = 3, "(4) $75,000-$99,999" = 4, "(5) $100,000 or more" = 5, "(99) DON’T KNOW/SKIPPED/REFUSED (VOL)" = 99)
data$SEX <- recode(data$SEX, "(1) Men" = 1, "(2) Women" =2, "(99) DON’T KNOW/SKIPPED/REFUSED (VOL)" = 99)
data$AGE <- recode(data$AGE65, "(1) 18-24" = 1, "(2) 25-29" = 2, "(3) 30-39" = 3, "(4) 40-49" = 4, "(5) 50-64" = 5, "(6) 65+" = 6, "(99) DON’T KNOW/SKIPPED/REFUSED (VOL)" = 99)
data$AGE65 <- NULL
data$PARTY <- recode(data$PARTYFULL, "(1) Democrat/Lean Dem" = 1, "(2) Republican/Lean Rep" =2, "(3) Independent" = 3)
data$PARTYFULL <- NULL
data$IDEO <- recode(data$IDEO, "(1) Very liberal" = 1, "(2) Somewhat liberal" = 2, "(3) Moderate" = 3, "(4) Somewhat conservative" = 4, "(5) Very conservative" = 5, "(99) DON’T KNOW/SKIPPED/REFUSED (VOL)" = 99)
data$RELIG <- recode(data$RELIG4, "(1) Protestant/Other Christian" = 1, "(2) Catholic"= 2, "(3) Other" =3, "(4) Unaffiliated" = 4, "(99) DON’T KNOW/SKIPPED/REFUSED (VOL)" = 99)
data$RELIG4 <- NULL
data$PLACE <- recode(data$SIZEPLACE, "(1) Urban" = 1, "(2) Suburban" = 2, "(3) Small town" = 3, "(4) Rural" = 4, "(99) DON’T KNOW/SKIPPED/REFUSED (VOL)" = 99)
data$SIZEPLACE <- NULL
data$TRACK <- recode(data$TRACK, "(1) Right direction" = 1, "(2) Wrong direction" = 2, "(99) DON’T KNOW/SKIPPED/REFUSED (VOL)" = 99)
data$APP <- recode(data$APP, "(1) Approve strongly" = 1, "(2) Approve somewhat" = 2, "(3) Disapprove somewhat" = 3, "(4) Disapprove strongly" = 4, "(99) DON’T KNOW/SKIPPED/REFUSED (VOL)" = 99)
data$STATE <- gsub("\\ .*","",data$STATE)
data$STATE <- substring(data$STATE, 2)
data$STATE <- gsub("\\).*","",data$STATE)
data$SU_ID <- NULL
data$RACE <- data$RACETH5
data$RACETH5 <- NULL
data$PLACE <- NULL
data$APP <- NULL
data$TRACK <- NULL
```

```{r}
data[1:5,]
```

```{r}
data <- filter_all(data, all_vars(. < 60))
```

```{r}
IDEO_data <- c()
for (i in 1:5){
  IDEO_data[i] <- sum(data$IDEO==i)
}
lbls <- c("(1) Very liberal", "(2) Somewhat liberal", "(3) Moderate", "(4) Somewhat conservative", "(5) Very conservative")
df <- data.frame(
  group = lbls,
  value = IDEO_data
  )
ggplot(df, aes(x="", y=value, fill=group))+ geom_bar(width = 1, stat = "identity")+ coord_polar("y", start=0)
```

```{r}
AGE_data <- c()
for (i in 1:6){
  AGE_data[i] <- sum(data$AGE==i)
}
lbls <- c("(1) 18-24",
"(2) 25-29",
"(3) 30-39",
"(4) 40-49",
"(5) 50-64",
"(6) 65+")
df <- data.frame(
  group = lbls,
  value = AGE_data
  )
ggplot(df, aes(x="", y=value, fill=group))+ geom_bar(width = 1, stat = "identity")+ coord_polar("y", start=0)
```

```{r}
data <- data %>% filter(EDUC<88)
EDUC_data <- c()
for (i in 1:4){
  EDUC_data[i] <- sum(data$EDUC==i)
}
lbls <- c("(1) High school or less",
"(2) Some college/assoc. degree",
"(3) College graduate",
"(4) Postgraduate study")
df <- data.frame(
  group = lbls,
  value = EDUC_data
  )
ggplot(df, aes(x="", y=value, fill=group))+ geom_bar(width = 1, stat = "identity")+ coord_polar("y", start=0)
```

```{r}
data <- data %>% filter(RACE<88)
RACE_data <- c()
for (i in 1:5){
  RACE_data[i] <- sum(data$RACE==i)
}
lbls <- c("(1) White",
"(2) African American or Black",
"(3) Latino or Hispanic",
"(4) Asian",
"(5) Other")
df <- data.frame(
  group = lbls,
  value = RACE_data
  )
ggplot(df, aes(x="", y=value, fill=group))+ geom_bar(width = 1, stat = "identity")+ coord_polar("y", start=0)
```

```{r}
data <- data %>% filter(PARTY<88)
PAR_data <- c()
for (i in 1:3){
  PAR_data[i] <- sum(data$PARTY==i)
}
lbls <- c("(1) Democrat/Lean Dem", "(2) Republican/Lean Rep", "(3) Independent")
df <- data.frame(
  group = lbls,
  value = PAR_data
  )
ggplot(df, aes(x="", y=value, fill=group))+ geom_bar(width = 1, stat = "identity")+ coord_polar("y", start=0)
```

```{r}
data <- data %>% filter(INCOME<88)
INC_data <- c()
for (i in 1:5){
  INC_data[i] <- sum(data$INCOME==i)
}
lbls <- c("(1) Under $25,000",
"(2) $25,000-$49,999",
"(3) $50,000-$74,999",
"(4) $75,000-$99,999",
"(5) $100,000 or more")
df <- data.frame(
  group = lbls,
  value = INC_data
  )
ggplot(df, aes(x="", y=value, fill=group))+ geom_bar(width = 1, stat = "identity")+ coord_polar("y", start=0)
```

```{r}
data <- data %>% filter(RELIG<88)
REL_data <- c()
for (i in 1:4){
  REL_data[i] <- sum(data$RELIG==i)
}
lbls <- c("(1) Protestant/Other Christian",
"(2) Catholic",
"(3) Other",
"(4) Unaffiliated")
df <- data.frame(
  group = lbls,
  value = REL_data
  )
ggplot(df, aes(x="", y=value, fill=group))+ geom_bar(width = 1, stat = "identity")+ coord_polar("y", start=0)
data$RELIG <- NULL
```

```{r}
table(data$IDEO, data$RACE)
table(data$IDEO, data$AGE)
table(data$IDEO, data$EDUC)
table(data$IDEO, data$INCOME)
```

```{r}
table(data$IDEO, data$STATE)
```

The group with the most conservative ideology is exposed to the greatest disclosure risk, across differenet demographic variables.

```{r}
table(data$PARTY, data$IDEO)
```

The model I want to use resembles the sequential model we learnt in the paper:
Age group $f(y_1)$, and follows a Dirichlet-multinomial prior $(p_{i1},p_{i2}, p_{i3},p_{i4},p_{i5}\sim Dirichlet(\alpha_1,\alpha_2,\alpha_3,\alpha_4,\alpha_5) $
Education level $f(y_2|y_1)$, and follows a multinomial logistic model with multiple explanatory variables (age groups) $log(\frac{p_{ic}}{p_{i1}})=\beta_{0c}+\beta_{1c}X_{age1}+\cdots+\beta_{5c}X_{age5}$
Income group $f(y_3|y_2,y_1)$, and is an ordered categorical variable
Ideology group $f(y_4|y_3,y_2,y_1)$, and follows a multinomial logistic model with multiple explanatory variables

**Synthesis Model**

First, I would use a binomial model to generate synthesized sex.

*Sex*

```{r}
modelString_sex <-"
model {
## sampling
for (i in 1:N){
y[i] ~ dbern(p)
}
## priors
p ~ dnorm(0.5, 1)
}
"
```

```{r}
data$SEX <- data$SEX - 1
y = as.vector(data$SEX)
N = length(y)  
```

```{r}
the_data <- list("y" = y,
                 "N" = N)
```

```{r}
initsfunction <- function(chain){
  .RNG.seed <- c(1,2)[chain]
  .RNG.name <- c("base::Super-Duper",
                 "base::Wichmann-Hill")[chain]
  return(list(.RNG.seed=.RNG.seed,
              .RNG.name=.RNG.name))
}
```

```{r}
posterior_sex <- run.jags(modelString_sex,
                      n.chains = 1,
                      data = the_data,
                      monitor = c("p"),
                      adapt = 1000,
                      burnin = 5000,
                      sample = 5000,
                      thin = 1,
                      inits = initsfunction)
```

```{r}
plot(posterior_sex, vars="p")
```

```{r}
post <- as.mcmc(posterior_sex)
```

```{r}
post
synthesize <- function(index, n){
  synthetic_sex <- c()
  synthetic_sex <- rbinom(n, 1, post[index,"p"])
  data.frame(synthetic_sex)
}
n <- dim(data)[1]
syn_sex <- data.frame(synthesize(1, n))
data <- cbind(data, syn_sex)
colnames(data)[10] <- "SEX_s"
```

Then, I will use a Dirichlet model to generate age and race.

*Age*

```{r eval=FALSE, include=TRUE}
modelString_age <-"
model {
## sampling
for (i in 1:N){
y[i,] ~ dmulti(p[i,1:C],1)
p[i,] ~ ddirch(alpha[])
}
## priors
for (c in 1:C){
alpha[c] <- 1
}
}
"
```

```{r eval=FALSE, include=TRUE} 
y = as.vector(data$AGE)
N = length(y)
C = 6
```

```{r eval=FALSE, include=TRUE}
the_data <- list("y" = y,
                 "N" = N,
                 "C" = C)
```

```{r eval=FALSE, include=TRUE}
posterior_sex <- run.jags(modelString_age,
                      n.chains = 1,
                      data = the_data,
                      monitor = c("p"),
                      adapt = 1000,
                      burnin = 5000,
                      sample = 5000,
                      thin = 1,
                      inits = initsfunction)
```

```{r eval=FALSE, include=TRUE}
post <- as.mcmc(posterior_age)
```

*Race*

```{r eval=FALSE, include=TRUE}
modelString_race <-"
model {
## sampling
for (i in 1:N){
y[i,] ~ dmulti(p[i,1:C],1)
p[i,] ~ ddirch(alpha[])
}
## priors
for (c in 1:C){
alpha[c] <- 1
}
}
"
```

```{r eval=FALSE, include=TRUE}
y = as.vector(data$RACE)
N = length(y)
C = 5
```

```{r eval=FALSE, include=TRUE}
the_data <- list("y" = y,
                 "N" = N,
                 "C" = C)
```

```{r eval=FALSE, include=TRUE}
posterior_sex <- run.jags(modelString_race,
                      n.chains = 1,
                      data = the_data,
                      monitor = c("p"),
                      adapt = 1000,
                      burnin = 5000,
                      sample = 5000,
                      thin = 1,
                      inits = initsfunction)
```

Moving on, I will use a multinomial logistic model to synthesize education and income.

```{r}
data$Age1 = fastDummies::dummy_cols(data$AGE)[,names(fastDummies::dummy_cols(data$AGE)) == ".data_1"]
data$Age2 = fastDummies::dummy_cols(data$AGE)[,names(fastDummies::dummy_cols(data$AGE)) == ".data_2"]
data$Age3 = fastDummies::dummy_cols(data$AGE)[,names(fastDummies::dummy_cols(data$AGE)) == ".data_3"]
data$Age4 = fastDummies::dummy_cols(data$AGE)[,names(fastDummies::dummy_cols(data$AGE)) == ".data_4"]
data$Age5 = fastDummies::dummy_cols(data$AGE)[,names(fastDummies::dummy_cols(data$AGE)) == ".data_5"]
data$Age6 = fastDummies::dummy_cols(data$AGE)[,names(fastDummies::dummy_cols(data$AGE)) == ".data_6"]
data$Race1 = fastDummies::dummy_cols(data$RACE)[,names(fastDummies::dummy_cols(data$RACE)) == ".data_1"]
data$Race2 = fastDummies::dummy_cols(data$RACE)[,names(fastDummies::dummy_cols(data$RACE)) == ".data_2"]
data$Race3 = fastDummies::dummy_cols(data$RACE)[,names(fastDummies::dummy_cols(data$RACE)) == ".data_3"]
data$Race4 = fastDummies::dummy_cols(data$RACE)[,names(fastDummies::dummy_cols(data$RACE)) == ".data_4"]
data$Race5 = fastDummies::dummy_cols(data$RACE)[,names(fastDummies::dummy_cols(data$RACE)) == ".data_5"]
data$WOMAN <- data$SEX_s
```

*Education*
```{r message = FALSE, size = "footnotesize"}
modelString_educ <-"
model {
## sampling
for (i in 1:N){
y[i] ~ dnorm(beta0 + beta1*Age1[i] + beta2*Age2[i] + beta3*Age3[i] + beta4*Age4[i] + beta5*Age5[i] + beta6*Race1[i] + beta7*Race2[i] + beta8*Race3[i] + beta9*Race4[i] + beta10*Woman[i], invsigma2)
}
## priors
beta0 ~ dnorm(0, 0.00001)
beta1 ~ dnorm(0, 0.00001)
beta2 ~ dnorm(0, 0.00001)
beta3 ~ dnorm(0, 0.00001)
beta4 ~ dnorm(0, 0.00001)
beta5 ~ dnorm(0, 0.00001)
beta6 ~ dnorm(0, 0.00001)
beta7 ~ dnorm(0, 0.00001)
beta8 ~ dnorm(0, 0.00001)
beta9 ~ dnorm(0, 0.00001)
beta10 ~ dnorm(0, 0.00001)
invsigma2 ~ dgamma(a, b)
sigma <- sqrt(pow(invsigma2, -1))
}
"
```

```{r}
y = as.vector(data$EDUC)
Age1 = as.vector(data$Age1)
Age2 = as.vector(data$Age2)
Age3 = as.vector(data$Age3)
Age4 = as.vector(data$Age4)
Age5 = as.vector(data$Age5)
Race1 = as.vector(data$Race1) 
Race2 = as.vector(data$Race2) 
Race3 = as.vector(data$Race3) 
Race4 = as.vector(data$Race4) 
Woman = as.vector(data$SEX_s)
N = length(y) 
```

```{r message = FALSE, size = "footnotesize"}
the_data <- list("y" = y,
                 "N" = N,
                 "Age1" = Age1,
                 "Age2" = Age2,
                 "Age3" = Age3,
                 "Age4" = Age4,
                 "Age5" = Age5,
                 "Race1" = Race1,
                 "Race2" = Race2,
                 "Race3" = Race3,
                 "Race4" = Race4,
                 "Woman" = Woman,
                 "a" = 1, "b" = 1)
```

```{r eval=FALSE, message=FALSE, include=FALSE, results=, size="footnotesize"}
posterior <- run.jags(modelString_educ,
                      n.chains = 1,
                      data = the_data,
                      monitor = c("beta0", "beta1", "beta2", "beta3", "beta4", "beta5", "beta6", "beta7", "beta8", "beta9", "beta10", "sigma"),
                      adapt = 1000, 
                      burnin = 5000,
                      sample = 5000,
                      thin = 1,
                      inits = initsfunction)
```

```{r message=FALSE, size = "footnotesize"}
post <- as.mcmc(posterior)
```

```{r message=FALSE, size = "footnotesize"}
synf_educ <- function(Age1, Age2, Age3, Age4, Age5, Race1, Race2, Race3, Race4, Woman, index, n){
  mean_Y <- post[index, "beta0"] +  Age1 * post[index, "beta1"] + Age2 * post[index, "beta2"] + Age3 * post[index, "beta3"] + Age4 * post[index, "beta4"] + Age5 * post[index, "beta5"] + Race1 * post[index, "beta6"] + Race1 * post[index, "beta7"] + Race1 * post[index, "beta8"] + Race1 * post[index, "beta9"] + Woman * post[index, "beta10"]
  synthetic_educ <- rnorm(n, mean_Y, post[index, "sigma"])
  data.frame(synthetic_educ)
}
index=1
n <- dim(data)[1]
syn_educ <- data.frame(synf_educ(data$Age1, data$Age2, data$Age3, data$Age4, data$Age5, data$Race1, data$Race2, data$Race3, data$Race4, data$SEX_s, 1, n))
syn_educ <- round(syn_educ$synthetic_educ,0)
data <- cbind(data, syn_educ)
colnames(data)[23] <- "EDUC_s"
data <- data %>% filter(EDUC_s >=1, EDUC_s <=4)
```

*Income*

```{r}
data$Educ1 = fastDummies::dummy_cols(data$EDUC_s)[,names(fastDummies::dummy_cols(data$EDUC_s)) == ".data_1"]
data$Educ2 = fastDummies::dummy_cols(data$EDUC_s)[,names(fastDummies::dummy_cols(data$EDUC_s)) == ".data_2"]
data$Educ3 = fastDummies::dummy_cols(data$EDUC_s)[,names(fastDummies::dummy_cols(data$EDUC_s)) == ".data_3"]
```

```{r message = FALSE, size = "footnotesize"}
modelString_inc <-"
model {
## sampling
for (i in 1:N){
y[i] ~ dnorm(beta0 + beta1*Age1[i] + beta2*Age2[i] + beta3*Age3[i] + beta4*Age4[i] + beta5*Age5[i] + beta6*Race1[i] + beta7*Race2[i] + beta8*Race3[i] + beta9*Race4[i] + beta10*Woman[i] + beta11*Educ1[i] + beta12*Educ2[i] + beta13*Educ3[i], invsigma2)
}
## priors
beta0 ~ dnorm(0, 0.00001)
beta1 ~ dnorm(0, 0.00001)
beta2 ~ dnorm(0, 0.00001)
beta3 ~ dnorm(0, 0.00001)
beta4 ~ dnorm(0, 0.00001)
beta5 ~ dnorm(0, 0.00001)
beta6 ~ dnorm(0, 0.00001)
beta7 ~ dnorm(0, 0.00001)
beta8 ~ dnorm(0, 0.00001)
beta9 ~ dnorm(0, 0.00001)
beta10 ~ dnorm(0, 0.00001)
beta11 ~ dnorm(0, 0.00001)
beta12 ~ dnorm(0, 0.00001)
beta13 ~ dnorm(0, 0.00001)
invsigma2 ~ dgamma(a, b)
sigma <- sqrt(pow(invsigma2, -1))
}
"
```

```{r}
y = as.vector(data$EDUC)
Age1 = as.vector(data$Age1)
Age2 = as.vector(data$Age2)
Age3 = as.vector(data$Age3)
Age4 = as.vector(data$Age4)
Age5 = as.vector(data$Age5)
Race1 = as.vector(data$Race1) 
Race2 = as.vector(data$Race2) 
Race3 = as.vector(data$Race3) 
Race4 = as.vector(data$Race4) 
Woman = as.vector(data$SEX_s)
Educ1 = as.vector(data$Educ1)
Educ2 = as.vector(data$Educ2)
Educ3 = as.vector(data$Educ3)
N = length(y) 
```

```{r message = FALSE, size = "footnotesize"}
the_data <- list("y" = y,
                 "N" = N,
                 "Age1" = Age1,
                 "Age2" = Age2,
                 "Age3" = Age3,
                 "Age4" = Age4,
                 "Age5" = Age5,
                 "Race1" = Race1,
                 "Race2" = Race2,
                 "Race3" = Race3,
                 "Race4" = Race4,
                 "Woman" = Woman,
                 "Educ1" = Educ1,
                 "Educ2" = Educ2,
                 "Educ3" = Educ3,
                 "a" = 1, "b" = 1)
```

```{r eval=FALSE, message=FALSE, include=FALSE, results=, size="footnotesize"}
posterior <- run.jags(modelString_inc,
                      n.chains = 1,
                      data = the_data,
                      monitor = c("beta0", "beta1", "beta2", "beta3", "beta4", "beta5", "beta6", "beta7", "beta8", "beta9", "beta10", "beta11", "beta12", "beta13", "sigma"),
                      adapt = 1000, 
                      burnin = 5000,
                      sample = 5000,
                      thin = 1,
                      inits = initsfunction)
```

```{r message=FALSE, size = "footnotesize"}
post <- as.mcmc(posterior)
```

```{r message=FALSE, size = "footnotesize"}
synf_inc <- function(Age1, Age2, Age3, Age4, Age5, Race1, Race2, Race3, Race4, Woman,Educ1, Educ2, Educ3, index, n){
  mean_Y <- post[index, "beta0"] +  Age1 * post[index, "beta1"] + Age2 * post[index, "beta2"] + Age3 * post[index, "beta3"] + Age4 * post[index, "beta4"] + Age5 * post[index, "beta5"] + Race1 * post[index, "beta6"] + Race1 * post[index, "beta7"] + Race1 * post[index, "beta8"] + Race1 * post[index, "beta9"] + Woman * post[index, "beta10"] + Educ1 * post[index, "beta11"] + Educ2 * post[index, "beta12"] + Educ3 * post[index, "beta13"]
  synthetic_inc <- rnorm(n, mean_Y, post[index, "sigma"])
  data.frame(synthetic_inc)
}
n <- dim(data)[1]
syn_inc <- data.frame(synf_inc(data$Age1, data$Age2, data$Age3, data$Age4, data$Age5, data$Race1, data$Race2, data$Race3, data$Race4, data$SEX_s, data$Educ1, data$Educ2, data$Educ3, 1, n))
syn_inc <- round(syn_inc$synthetic_inc,0)
data <- cbind(data, syn_inc)
colnames(data)[27] <- "INC_s"
data <- data %>% filter(INC_s >=1, INC_s <=5)
```

Finally! I will use age, race, sex, education, and income to synthesize ideology.

```{r}
data$Inc1 = fastDummies::dummy_cols(data$INC_s)[,names(fastDummies::dummy_cols(data$INC_s)) == ".data_1"]
data$Inc2 = fastDummies::dummy_cols(data$INC_s)[,names(fastDummies::dummy_cols(data$INC_s)) == ".data_2"]
data$Inc3 = fastDummies::dummy_cols(data$INC_s)[,names(fastDummies::dummy_cols(data$INC_s)) == ".data_3"]
data$Inc4 = fastDummies::dummy_cols(data$INC_s)[,names(fastDummies::dummy_cols(data$INC_s)) == ".data_4"]
```

```{r message = FALSE, size = "footnotesize"}
modelString_ideo <-"
model {
## sampling
for (i in 1:N){
y[i] ~ dnorm(beta0 + beta1*Age1[i] + beta2*Age2[i] + beta3*Age3[i] + beta4*Age4[i] + beta5*Age5[i] + beta6*Race1[i] + beta7*Race2[i] + beta8*Race3[i] + beta9*Race4[i] + beta10*Woman[i] + beta11*Educ1[i] + beta12*Educ2[i] + beta13*Educ3[i] + beta14*Inc1[i] + beta15*Inc2[i] + beta16*Inc3[i] + beta17*Inc4[i], invsigma2)
}
## priors
beta0 ~ dnorm(0, 0.00001)
beta1 ~ dnorm(0, 0.00001)
beta2 ~ dnorm(0, 0.00001)
beta3 ~ dnorm(0, 0.00001)
beta4 ~ dnorm(0, 0.00001)
beta5 ~ dnorm(0, 0.00001)
beta6 ~ dnorm(0, 0.00001)
beta7 ~ dnorm(0, 0.00001)
beta8 ~ dnorm(0, 0.00001)
beta9 ~ dnorm(0, 0.00001)
beta10 ~ dnorm(0, 0.00001)
beta11 ~ dnorm(0, 0.00001)
beta12 ~ dnorm(0, 0.00001)
beta13 ~ dnorm(0, 0.00001)
beta14 ~ dnorm(0, 0.00001)
beta15 ~ dnorm(0, 0.00001)
beta16 ~ dnorm(0, 0.00001)
beta17 ~ dnorm(0, 0.00001)
invsigma2 ~ dgamma(a, b)
sigma <- sqrt(pow(invsigma2, -1))
}
"
```

```{r}
y = as.vector(data$EDUC)
Age1 = as.vector(data$Age1)
Age2 = as.vector(data$Age2)
Age3 = as.vector(data$Age3)
Age4 = as.vector(data$Age4)
Age5 = as.vector(data$Age5)
Race1 = as.vector(data$Race1) 
Race2 = as.vector(data$Race2) 
Race3 = as.vector(data$Race3) 
Race4 = as.vector(data$Race4) 
Woman = as.vector(data$SEX_s)
Educ1 = as.vector(data$Educ1)
Educ2 = as.vector(data$Educ2)
Educ3 = as.vector(data$Educ3)
Inc1 = as.vector(data$Inc1)
Inc2 = as.vector(data$Inc2)
Inc3 = as.vector(data$Inc3)
Inc4 = as.vector(data$Inc4)
N = length(y) 
```

```{r message = FALSE, size = "footnotesize"}
the_data <- list("y" = y,
                 "N" = N,
                 "Age1" = Age1,
                 "Age2" = Age2,
                 "Age3" = Age3,
                 "Age4" = Age4,
                 "Age5" = Age5,
                 "Race1" = Race1,
                 "Race2" = Race2,
                 "Race3" = Race3,
                 "Race4" = Race4,
                 "Woman" = Woman,
                 "Educ1" = Educ1,
                 "Educ2" = Educ2,
                 "Educ3" = Educ3,
                 "Inc1" = Inc1,
                 "Inc2" = Inc2,
                 "Inc3" = Inc3,
                 "Inc4" = Inc4,
                 "a" = 1, "b" = 1)
```

```{r eval=FALSE, message=FALSE, include=FALSE, results=, size="footnotesize"}
posterior <- run.jags(modelString_ideo,
                      n.chains = 1,
                      data = the_data,
                      monitor = c("beta0", "beta1", "beta2", "beta3", "beta4", "beta5", "beta6", "beta7", "beta8", "beta9", "beta10", "beta11", "beta12", "beta13", "beta14", "beta15", "beta16", "beta17","sigma"),
                      adapt = 1000, 
                      burnin = 5000,
                      sample = 5000,
                      thin = 1,
                      inits = initsfunction)
```

```{r message=FALSE, size = "footnotesize"}
post <- as.mcmc(posterior)
```

```{r message=FALSE, size = "footnotesize"}
synf_ideo <- function(Age1, Age2, Age3, Age4, Age5, Race1, Race2, Race3, Race4, Woman,Educ1, Educ2, Educ3, Inc1, Inc2, Inc3, Inc4, index, n){
  mean_Y <- post[index, "beta0"] +  Age1 * post[index, "beta1"] + Age2 * post[index, "beta2"] + Age3 * post[index, "beta3"] + Age4 * post[index, "beta4"] + Age5 * post[index, "beta5"] + Race1 * post[index, "beta6"] + Race1 * post[index, "beta7"] + Race1 * post[index, "beta8"] + Race1 * post[index, "beta9"] + Woman * post[index, "beta10"] + Educ1 * post[index, "beta11"] + Educ2 * post[index, "beta12"] + Educ3 * post[index, "beta13"] + Inc1 * post[index, "beta14"] + Inc2 * post[index, "beta15"] + Inc3 * post[index, "beta16"] + Inc4 * post[index, "beta17"]
  synthetic_ideo <- rnorm(n, mean_Y, post[index, "sigma"])
  data.frame(synthetic_ideo)
}
n <- dim(data)[1]
syn_ideo <- data.frame(synf_ideo(data$Age1, data$Age2, data$Age3, data$Age4, data$Age5, data$Race1, data$Race2, data$Race3, data$Race4, data$SEX_s, data$Educ1, data$Educ2, data$Educ3, data$Inc1, data$Inc2, data$Inc3, data$Inc4, 1, n))
syn_ideo <- round(syn_ideo$synthetic_ideo,0)
data <- cbind(data, syn_ideo)
colnames(data)[32] <- "IDEO_s"
data <- data %>% filter(IDEO_s >=1, IDEO_s <=5)
```

```{r}
test <- cbind(data$STATE, data$SEX, data$EDUC, data$INCOME, data$IDEO, data$AGE, data$RACE, data$SEX_s, data$EDUC_s, data$INC_s, data$IDEO_s)
test <- data.frame(test)
names(test) <- c("STATE", "SEX_o", "EDUC_o", "INCOME_o", "IDEO_o", "AGE_o", "RACE_o", "SEX_s", "EDUC_s", "INC_s", "IDEO_s")
test
```

```{r}
IDEO_o <- c()
for (i in 1:5){
  IDEO_o[i] <- sum(test$IDEO_o==i)
}
lbls <- c("(1) Very liberal", "(2) Somewhat liberal", "(3) Moderate", "(4) Somewhat conservative", "(5) Very conservative")
df <- data.frame(
  group = lbls,
  value = IDEO_o
  )
ggplot(df, aes(x="", y=value, fill=group))+ geom_bar(width = 1, stat = "identity")+ coord_polar("y", start=0)
```

```{r}
IDEO_s <- c()
for (i in 1:5){
  IDEO_s[i] <- sum(test$IDEO_s==i)
}
lbls <- c("(1) Very liberal", "(2) Somewhat liberal", "(3) Moderate", "(4) Somewhat conservative", "(5) Very conservative")
df <- data.frame(
  group = lbls,
  value = IDEO_s
  )
ggplot(df, aes(x="", y=value, fill=group))+ geom_bar(width = 1, stat = "identity")+ coord_polar("y", start=0)
```

```{r}
mean(as.numeric(test$IDEO_o))
mean(as.numeric(test$IDEO_s))
median(as.numeric(test$IDEO_o))
median(as.numeric(test$IDEO_s))
```

We could see that the moderate groups are better preserved in our synthesized datasets, whereas the groups with more radical ideologies got shrunk a little bit. 

```{r}
test
```


```{r}
model_o <- lm(as.numeric(IDEO_o) ~ SEX_o + EDUC_o + INCOME_o, data=test)
model_s <- lm(as.numeric(IDEO_s) ~ SEX_o + EDUC_o + INCOME_o, data=test)
```

```{r}
summary(model_o)
```

```{r}
summary(model_s)
```

Judging from R-squared, we could see that my synthesized data has lost some utility in terms of the relationship between ideology and other variables in the dataset.
