---
title: "ACS Data Identification Disclosure Risk"
author: "Henrik Olsson"
date: "March 24, 2020"
output:
  pdf_document: default
  word_document: default
  html_document:
    df_print: paged
subtitle: MATH 301 Data Confidentiality
---

```{r include=FALSE}
library(readr)
library(LearnBayes)
library(plyr)
library(dplyr)
library(ggplot2)
library(runjags)
library(coda)
library(imputeTS)
library(fastDummies)
library(tinytex)
```

```{r}
ACSdata_org <- read.csv("ACSdata_org.csv")
ACSdata_syn <- read.csv("ACSdata_syn.csv")
ACSdata_syn2 <- read.csv("ACSdata_syn2.csv")
ACSdata_syn3 <- read.csv("ACSdata_syn3.csv")
```


Note: The datasets exlucde the HISP variable in the ACS data dictionary, but every other variable stays the same with the same name and description.

The following four variables are synthesized: LANX, WAOB, DIS, HICOV.


### Where m = 3 case 

#### Step 1: calculate key quantities

```{r echo = FALSE}
## Step 1: calculate key quantities 

CalculateKeyQuantities <- function(origdata, syndata, known.vars, syn.vars, n){
  origdata <- origdata
  syndata <- syndata
  n <- n
  c_vector <- rep(NA, n)
  T_vector <- rep(NA, n)
  for (i in 1:n){
    match <- (eval(parse(text=paste("origdata$",syn.vars,"[i]==
                                      syndata$",syn.vars,sep="",collapse="&")))&
                  eval(parse(text=paste("origdata$",known.vars,"[i]==
                                        syndata$",known.vars,sep="",collapse="&"))))
    match.prob <- ifelse(match, 1/sum(match), 0)
    
    if (max(match.prob) > 0){
      c_vector[i] <- length(match.prob[match.prob == max(match.prob)])
    }
    else
      c_vector[i] <- 0
      T_vector[i] <- is.element(i, rownames(origdata)[match.prob == max(match.prob)])
  }
  
  K_vector <- (c_vector * T_vector == 1)
  F_vector <- (c_vector * (1 - T_vector) == 1)
  s <- length(c_vector[c_vector == 1 & is.na(c_vector) == FALSE])
  
  res_r <- list(c_vector = c_vector,
                T_vector = T_vector,
                K_vector = K_vector,
                F_vector = F_vector,
                s = s
  )
  return(res_r)
}
```


- four synthesized variables: LANX, WAOB, DIS, HICOV, assigned to syn.vars
- three known variables: SEX, RACE, MAR, assigned to known.vars

```{r}
known.vars <- c("SEX", "RACE", "MAR")
syn.vars <- c("LANX", "WAOB", "DIS", "HICOV")
n <- dim(ACSdata_org)[1]
KeyQuantities1 <- CalculateKeyQuantities(ACSdata_org, ACSdata_syn, known.vars, syn.vars, n)
KeyQuantities2 <- CalculateKeyQuantities(ACSdata_org, ACSdata_syn2, known.vars, syn.vars, n)
KeyQuantities3 <- CalculateKeyQuantities(ACSdata_org, ACSdata_syn3, known.vars, syn.vars, n)

## Step 2: calculate 3 summary measures 

IdentificationRisk <- function(c_vector, T_vector, K_vector, F_vector, s, N){
  
  nonzero_c_index <- which(c_vector > 0)
  exp_match_risk <- sum(1/c_vector[nonzero_c_index]*T_vector[nonzero_c_index])
  true_match_rate <- sum(na.omit(K_vector))/N
  false_match_rate <- sum(na.omit(F_vector))/s
  res_r <- list(exp_match_risk = exp_match_risk,
                true_match_rate = true_match_rate,
                false_match_rate = false_match_rate
  )
  return(res_r)
}

## each record is a target, therefore N = n

c_vector1 <- KeyQuantities1[["c_vector"]]
T_vector1 <- KeyQuantities1[["T_vector"]]
K_vector1 <- KeyQuantities1[["K_vector"]]
F_vector1 <- KeyQuantities1[["F_vector"]]
s1 <- KeyQuantities1[["s"]]
N <- n
ThreeSummaries1 <- IdentificationRisk(c_vector1, T_vector1, K_vector1, F_vector1, s, N)

c_vector2 <- KeyQuantities2[["c_vector"]]
T_vector2 <- KeyQuantities2[["T_vector"]]
K_vector2 <- KeyQuantities2[["K_vector"]]
F_vector2 <- KeyQuantities2[["F_vector"]]
s2 <- KeyQuantities2[["s"]]
N <- n
ThreeSummaries2 <- IdentificationRisk(c_vector2, T_vector2, K_vector2, F_vector2, s2, N)

c_vector3 <- KeyQuantities3[["c_vector"]]
T_vector3 <- KeyQuantities3[["T_vector"]]
K_vector3 <- KeyQuantities3[["K_vector"]]
F_vector3 <- KeyQuantities3[["F_vector"]]
s3 <- KeyQuantities3[["s"]]
N <- n
ThreeSummaries3 <- IdentificationRisk(c_vector3, T_vector3, K_vector3, F_vector3, s3, N)
```

##### Summaries:

```{r}
## Expected match risk
(ThreeSummaries1[["exp_match_risk"]] + ThreeSummaries2[["exp_match_risk"]] + ThreeSummaries3[["exp_match_risk"]])/3
```

```{r}
## True match rate
(ThreeSummaries1[["true_match_rate"]] + ThreeSummaries2[["true_match_rate"]] + ThreeSummaries3[["true_match_rate"]])/3
```

```{r}
## False match rate
(ThreeSummaries3[["false_match_rate"]] + ThreeSummaries3[["false_match_rate"]] + ThreeSummaries3[["false_match_rate"]])/3
```

### Results and Discussion

```{r}
41.46743/10000
```

The expected match risk is 0.00414 probability average for each record to be correctly identified.

```{r}
0.0005666667*10000
```

5.666667, or approximately 6 records are correct unique matches.

```{r}
(s1 * ThreeSummaries3[["false_match_rate"]] + s2 * ThreeSummaries3[["false_match_rate"]] + s3 * ThreeSummaries3[["false_match_rate"]])/3
```

The 0.964 false match rate: among the 195 (the value of each s1, s2, s3) unique matches, 155 are false matches.



