---
title: "Identification Disclosure Risk- Hu Paper"
author: "Henrik Olsson"
date: "March 1, 2020"
output:
  pdf_document: default
  word_document: default
  html_document:
    df_print: paged
subtitle: MATH 301 Data Confidentiality
---

```{r include=FALSE}
library(readr)
library(LearnBayes)
library(plyr)
library(dplyr)
library(ggplot2)
library(runjags)
library(coda)
library(imputeTS)
library(fastDummies)
library(tinytex)
```

```{r}
acsdata_org <- read.csv("ACSdata_org.csv")
acsdata_syn <- read.csv("ACSdata_syn.csv")
```


Let $c_i$ be the number of records with the highest match probability for the target $t_i$; let $T_i = 1$ if the true match is among the $c_i$ units and $T_i = 0$ otherwise. Let $K_i = 1$ when $c_iT_i = 1$ and $K_i = 0$ otherwise, and let $N$ denote the total number of target records. Finally, let $F_i = 1$ when $c_i(1-T_i) = 1$ and $F_i = 0$ otherwise, and let $s$ equal the number of records with $c_i = 1$.

### (i) The expected match risk:



```{r eval=FALSE, include=FALSE}
total_data <- cbind(acsdata_org, syn_sex, syn_race, syn_mar)

expect_match = 0
true_match = 0
N = dim(acsdata_org)

for(i in 1:N){
  select <- which(acsdata_org$SEX == syn_sex[i] & acsdata_org$MAR == syn_mar[i] & acsdata_org$RACE == syn_race[i])
  c_i <- length(select)
  if(i == select){
    expected_match_risk = expect_match + 1/c_i
  }
}
```





```{r}
## Reference from Kevin Ros
syn_sex <- acsdata_syn$SEX 
syn_race <- acsdata_syn$RACE
syn_mar <- acsdata_syn$MAR

N = nrow(acsdata_syn)
n = nrow(acsdata_org)

expected_match_risk = 0
true_match_rate = 0

for(i in 1:N){
  c_i = 0
  data = acsdata_org[acsdata_org$SEX == syn_sex[i] & acsdata_org$RACE == syn_race[i] & acsdata_org$MAR == syn_mar[i],]
  
  if(nrow(data) != 0){
    expected_match_risk = expected_match_risk + (1/nrow(data))
  }
  if(nrow(data) == 1)
    true_match_rate = true_match_rate + 1/N
}
print(expected_match_risk)
```


### (ii) The true match rate:

```{r}
print(true_match_rate)

```


### (iii) The false match rate:

```{r}
false_match_rate = 0
for(i in 1:N){
  
  c_i = 1
  s=nrow(c_i)
  data = acsdata_org[acsdata_org$SEX == syn_sex[i] & acsdata_org$RACE == syn_race[i] & acsdata_org$MAR == syn_mar[i],]
  
  if(nrow(data) == 1){
    false_match_rate = sum(false_match_rate/s) 
  }
}
print(false_match_rate)
```

